Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-11-07 boutiquespaper.tex
Updates following comments from Charles and Guillaume

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
\affil[2]{McGill University, Montreal, Canada}
\affil[3]{Montreal Neurological Institute, Montreal, Canada}
\affil[4]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montréal, QC, Canada}
\affil[5]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France}
\affil[6]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}
\affil[7]{Wellcome Trust Centre for Neuroimaging}
\affil[8]{Department of Psychology, Stanford University, Stanford, California, USA}
\affil[9]{Partners MS Center, Brigham and Women's Hospital, Brookline, Massachusetts, USA}
\affil[10]{Compute Canada}

@@

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are corralled in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform or replacing
existing functionality already represented in the \boutiques schema.

@@

\subsubsection{SPINE}
SPINE (\url{http://spinevirtuallab.org}), which stands for Structured
Planning and Implementation of New Explorations, is a web-based,
collaborative platform (virtual laboratory) designed to support the
design and execution of experiments centered on specific scientific
questions.  SPINE enables distributed data collection and management,
as well as experiment design, execution, and review. Boutiques serves
as SPINE's algorithm and workflow repository, and enables unequivocal
referencing of specific workflows applied to specified datasets within
an experiment, thereby describing the provenance and facilitating the
reproducibility of image-derived measurements. Workflows in SPINE may
combine human image annotation with automated image processing
algorithms. Future development will focus on extending Boutiques
workflow descriptors to include the identification and
characterization of human operators and their specific historic
performance on the required tasks. SPINE is currently hosted at
Brigham \& Women's Hospital in Boston, and supports several
international projects.

@@

\paragraph{CreaPhase}
The CreaPhase phase-contrast simulator~\cite{WEBE-17}, developed at
Creatis, was integrated in VIP through \boutiques. The inputs had
certain particularities (some needed to be enclosed in simple quotes,
other were vectors of variable size enclosed in brackets) that
required the post-processing of the wrapper script generated by the
\boutiques importer.

@@

BIDS apps~\cite{gorgolewski2017bids}, an effort for the adoption of
the Brain Imaging Data Structure (BIDS) in common neuroimaging
pipelines, require a standardized set of input and output
parameters. We developed a tool as part of the \boutiques importer to
generate a descriptor for any such BIDS app. We validated this tool by
importing BIDS apps containing the Statistical Parameter Mapping
toolbox (SPM)~\cite{penny2011statistical} and the ndmg pipeline
mentioned above. Descriptors are on Neurolinks.

@@

In addition, important runtime parameters, for instance related to
multi-threading or available resources (storage, RAM), may be set by
the execution platform without being specified in the \boutiques
descriptor. Such runtime parameters may impact reproducibility in some
cases. To properly cover this issue, \boutiques descriptors should be
complemented by a provenance framework that captures a detailed trace
of the execution. We plan to leverage the provenance format being
defined by the Neuro-Imaging Data Model
Workflow~\cite{ghosh2017capturing} initiative for this purpose.

>>>>

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
\affil[2]{McGill University, Montreal, Canada}
\affil[3]{Montreal Neurological Institute, Montreal, Canada}
\affil[4]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montréal, QC, Canada}
\affil[5]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France}
\affil[6]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}
\affil[7]{Wellcome Trust Centre for Neuroimaging, London, UK}
\affil[8]{Department of Psychology, Stanford University, Stanford, California, USA}
\affil[9]{Center for Neurological Imaging, Department of Radiology, Brigham and Women's Hospital, Boston, Massachusetts, USA}
\affil[10]{Compute Canada}

@@

Custom properties may be added to the Boutiques specification without
restriction. Custom properties are grouped together in a specific JSON object
to facilitate validation. They may be useful to implement
platform-specific features but they should be used with care to avoid
making applications dependent on a particular platform or replacing
existing functionality already represented in the \boutiques schema.

@@

\subsubsection{SPINE}
SPINE (\url{http://spinevirtuallab.org}), which stands for Structured
Planning and Implementation of New Explorations, is a web-based,
collaborative platform (virtual laboratory) designed to support the
design and execution of experiments centered on specific scientific
questions.  SPINE enables distributed data collection and management,
as well as experiment design, execution, and review. Boutiques will serve
as SPINE's algorithm and workflow repository, and enables unequivocal
referencing of specific workflows applied to specified datasets within
an experiment, thereby describing the provenance and facilitating the
reproducibility of image-derived measurements. Workflows in SPINE may
combine human image annotation with automated image processing
algorithms. Future development will focus on extending Boutiques
workflow descriptors to include the identification and
characterization of human operators and their specific historic
performance on the required tasks. SPINE is currently hosted at
Brigham \& Women's Hospital in Boston, and supports several
international projects.

@@

\paragraph{CreaPhase}
The CreaPhase phase-contrast simulator, developed at
Creatis, was integrated in VIP through \boutiques. The inputs had
certain particularities (some needed to be enclosed in simple quotes,
other were vectors of variable size enclosed in brackets) that
required the post-processing of the wrapper script generated by the
\boutiques importer.

@@

BIDS apps~\cite{gorgolewski2017bids}, an effort for the adoption of
the Brain Imaging Data Structure (BIDS) in common neuroimaging
pipelines, require a standardized set of input and output
parameters. We developed a tool as part of the \boutiques importer to
generate a descriptor for any such BIDS app. We validated this tool by
importing BIDS apps containing the Statistical Parametric Mapping
toolbox (SPM)~\cite{penny2011statistical} and the ndmg pipeline
mentioned above. Descriptors are on Neurolinks.

@@

In addition, important runtime parameters, for instance related to
multi-threading or available resources (storage, RAM), may be set by
the execution platform without being specified in the \boutiques
descriptor. Such runtime parameters may impact reproducibility in some
cases. To properly cover this issue, \boutiques descriptors should be
complemented by a provenance framework that captures a detailed trace
of the execution. We plan to leverage the provenance format being
defined by the NeuroImaging Data
Model-Workflow~\cite{ghosh2017capturing} initiative for this purpose.
